Master's Project Journal

### 2019-04-15 thesis/NetworkImplemenation.tex
49

Figure \ref{fig:updown_ctr} depicts the implementation of this counter. Apart from the aforementioned capping of the measurement range implemented using a pair of comparators and a multiplexer, important to note is the bit width of the increment. Despite overflow being disallowed when the number is interpreted as a signed number, the addition of a 7 bit wide -1 to the accumulator value will cause overflow in the adder. This, however, is perfectly valid as an implementation of subtraction in two's complement. On the right side of the diagram the output register can be seen. Clocked using the \ac{FSM}s measurement interval complete signal, this register preserves the phase error until the termination of the following interval.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{../bad_waveform}
	\caption[\ac{RTL} diagram of up-down counter]{\ac{RTL} diagram of up-down counter.}
	\label{fig:updown_ctr}
\end{figure}
%TODO updown counter RTL diagram

@@

\subsubsection{\acl{DCO}}
\ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of an \ac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout. Also advantageous is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design. Taking the formula given in Chapter \ref{chap:3} and the inverter propagation time used by Vivado in post synthesis/implementation simulations of $315~\si{\pico\second}$, the period step can be computed:
\begin{align}
t_{step} &= \text{inverters per step}\times\text{inversions per cycle}\times\text{propagation delay} \\
t_{step} &= 2\times 2\times 315~\si{\pico\second} = 1.26~\si{\nano\second}
\end{align}
Using this period step, the number of inverters required to produce the signal of period $200~\si{\nano\second}$ can easily be computed:
\begin{equation}
\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
\end{equation}

Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss

A control code width of 5 bits was chosen as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count. To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$. The corresponding minimum and maximum frequencies then are:
\begin{align}
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\0.5\timest_{step}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.136~\si{\nano\second}} = 5.332~\si{\mega\hertz}
\end{align}

@@

This is achieved by assigning specific areas of the chip in which that module must lie, although these must be of a size approximately $40\%$ larger than the minimum space required in order to avoid having the router create a complex, delay intensive layout to fit the \ac{RO}. To achieve greater consistency, the implementation directive can be modified such that the router will attempt to use the minimum area that does not require complex routing. \texttt{congestion_spreadlogic_low} was used for this purpose in this project, although other options may obtain similar results. The other options available in Vivado can be viewed in the \textit{Vivado Design Suite User Guide - Implementation} \cite{impl_ug}.

With the hardware description of the \ac{RO} completed, testing of an instance of the \ac{RO} with a constant control code of zero was attempted, and it was discovered that the estimation of the centre frequency based on the propagation delay was off by $100$s of $\si{\kilo\hertz}$. The addition of 100 inverters was required to restore a centre of $5~\si{\mega\hertz}$. After the \ac{RO} was integrated into \ac{ADPLL} 2 the centre frequency was discovered to have changed once more. The minor variation caused by the now variable control code had changed the average propagation delay through the inverters once more, now requiring 373 inverters for $5~\si{\mega\hertz}$.

When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delievered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$. For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used. To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
%TODO possibily problematic in the last two paragraphs

@@

\section{Minor Variations}\label{section:minor_variations}
%TODO osc side

>>>>

Figure \ref{fig:updown_ctr} depicts the implementation of this counter. Apart from the aforementioned capping of the measurement range implemented using a pair of comparators and a multiplexer, important to note is the bit width of the increment. Despite overflow being disallowed when the number is interpreted as a signed number, the addition of a 7 bit wide -1 to the accumulator value will cause overflow in the adder. This, however, is perfectly valid as an implementation of subtraction in two's complement. On the right side of the diagram the output register can be seen. Clocked using the \ac{FSM}s measurement interval complete signal, this register preserves the phase error until the termination of the following interval.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../updown}
	\caption[\ac{RTL} diagram of up-down counter]{\ac{RTL} diagram of up-down counter.}
	\label{fig:updown_ctr}
\end{figure}

@@

\subsubsection{\acl{DCO}}
\ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of an \ac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout. Also advantageous is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design. Taking the formula given in Chapter \ref{chap:3} and the inverter propagation time used by Vivado in post synthesis/implementation simulations of $315~\si{\pico\second}$, the period step can be computed:
\begin{align}
t_{step} &= \text{inverters per step}\times\text{inversions per cycle}\times\text{propagation delay} \\
t_{step} &= 2\times 2\times 315~\si{\pico\second} = 1.26~\si{\nano\second}
\end{align}
Using this period step, the number of inverters required to produce the signal of period $200~\si{\nano\second}$ can easily be computed:
\begin{equation}
\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
\end{equation}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../ro_new}
	\caption[\acl{RO} RTL diagram]{\acl{RO} RTL diagram.}
	\label{fig:ro_impl}
\end{figure}
Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss

A control code width of 5 bits was chosen as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count. To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$. The corresponding minimum and maximum frequencies then are:
\begin{align}
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.136~\si{\nano\second}} = 5.332~\si{\mega\hertz}
\end{align}

@@

This is achieved by assigning specific areas of the chip in which that module must lie, although these must be of a size approximately $40\%$ larger than the minimum space required in order to avoid having the router create a complex, delay intensive layout to fit the \ac{RO}. To achieve greater consistency, the implementation directive can be modified such that the router will attempt to use the minimum area that does not require complex routing. \texttt{congestion\_spreadlogic\_low} was used for this purpose in this project, although other options may obtain similar results. The other options available in Vivado can be viewed in the \textit{Vivado Design Suite User Guide - Implementation} \cite{impl_ug}.

With the hardware description of the \ac{RO} completed, testing of an instance of the \ac{RO} with a constant control code of zero was attempted, and it was discovered that the estimation of the centre frequency based on the propagation delay was off by $100$s of $\si{\kilo\hertz}$. The addition of 100 inverters was required to restore a centre of $5~\si{\mega\hertz}$. After the \ac{RO} was integrated into \ac{ADPLL} 2 the centre frequency was discovered to have changed once more. The minor variation caused by the now variable control code had changed the average propagation delay through the inverters once more, now requiring 373 inverters for $5~\si{\mega\hertz}$.
%TODO possibily problematic in the last two paragraphs

When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delievered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$. For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used. To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.

@@

\section{Minor Variations}\label{section:minor_variations}
%TODO osc size
